mathml: Translate R expressions to MathML and LaTeX/MathJax

### 2022-12-14 Draft3.Rmd
table 1 table 2

`mathml` contains several basic elements used in everyday mathematical expressions. 
For example it is possible to print the function for the linear regression as:

```{r, results='asis'}
term <- quote(hat(y) == hat(b)["0"] + hat(b)["1"]*x)
mathout(term)

# or with greek letters as:
term <- quote(y == beta["0"] + beta["1"]*x + ... + varepsilon["i"])

@@

_I also wanted to add a condition like 1 < a < 2, but apparently it's not possible_ 
It is also possible to represent a system of equations with different relations. Note that you may also assign a name to the function. The default name is in fact the greek letter lambda, but with `attributes()$name` the name costumization is performed smoothly.

@@

mathout(f)

attributes(f)$name = quote(f)
mathout(f)

@@

## Subscripts and Superscripts
Secondly, \pkg{mathml} allows also for the display of subscripts, superscripts and their combination.

```{r, results="asis", echo=TRUE}
catmath(a["x"])

catmath(a^x)

catmath(x["b"]^"a")
```

## Fractions
As in [@murrell2000] the `frac` and `over` functions aim to represent fractions. `choose` and `lchoose`, as in R\ \pkg{base}, display the binomial coefficient.

```{r, results="asis", echo=TRUE}
catmath(frac(x, y))

catmath(over(x, y))

catmath(choose(n, k))

catmath(lchoose(n, k))
```

## Radicals

```{r, results="asis", echo=TRUE}
catmath(sqrt(x))
```

## Relations

Among the possible relational operators, following [@murrell2000], we included:

```{r, results="asis", echo=TRUE}
catmath(x==y)

catmath(x < y)

catmath(x <= y)

catmath(x >= y)

catmath(!x)

catmath(x != y)

catmath(X ~ Y)

catmath(x %~~% y)

catmath(x %==% y)

catmath(x %=~% y)

catmath(x %prop% y)
```

## Set relations

Here is provided a series of useful set relations.

_Still to be implemented_

```{r, results="asis", echo=TRUE}
catmath(a %in% A)

catmath(intersect(A, B))

catmath(union(A, B))

catmath(is.null(intersect(A, B)))
```

## Arrows

Following once again [@murrell2000], we provide here a list of useful arrows

```{r, results="asis", echo=TRUE}
# x %<->% y
catmath(x %<->% y)

# x %->% y
catmath(x %->% y)

# x %<-% y
catmath(x %<-% y)

# x %up% y:
catmath(x %up% y)

# x %down% y:
catmath(x %down% y)

# x %<=>% y: solved it using iff
catmath(x %<=>% y)

# x %=>% y: solved using rArr
catmath(x %=>% y)

# x %<=% y: solved using lArr
catmath(x %<=% y)

# x %dblup% y:
catmath(x %dblup% y)

# x %dbldown% y:
catmath(x %dbldown% y)
```

## Accents

```{r, results="asis", echo=TRUE}
catmath(hat(x))

catmath(hat(xy))

catmath(mean(x))

catmath(mean(xy))
```

## Symbolic names

```{r, results="asis", echo=TRUE}
catmath(A-Omega)

catmath(alpha-omega)

catmath(pi-Pi)

catmath(Inf)
```

@@

## Absolute value

Is it also possible to calculate and represent the absolute value, as:

```{r, results='asis', echo=TRUE}
catmath(abs(x))
```

@@

# A case study

>>>>

\pkg{mathml} contains several basic elements used in everyday mathematical expressions, such as superscripts, subscripts, hat, relation and set signs, accents and arrows. Below we provide some examples of their application with commented code. The other basic elements will be displayed in Table \@ref(tab:basic-elements) and Table \@ref(tab:arrows).

Let us start with the function for the linear regression. As the reader may notice, \pkg{mathml} displays Greek letters, accents and superscripts.

```{r, results='asis'}
term <- quote(hat(y) == hat(b)["0"] + hat(b)["1"]*x["1"])
mathout(term)

# or with greek letters as:
term <- quote(y == beta["0"] + beta["1"]*x["1"] + beta["2"]*x["2"] + ... + beta["i"]*x["i"] + varepsilon["i"])

@@

The second example may be particularly of use when we deal with multiple conditions (see \@ref(case-study)). Here, for instance, we present a trivial one with different relation signs.

@@

Note that you may also assign a name to the function. The default name is in fact the greek letter lambda, but with `attributes()$name` the name costumization is performed smoothly.

```{r, results='asis'}
attributes(f)$name = quote(f)
mathout(f)
```

```{r basic-elements, echo=FALSE}
op <- list(subsup=quote(x["a"]^b), lchoose=quote(lchoose(n, k)), sqrt=quote(sqrt(x)), not=quote(!x), different=quote(x != y), mean=quote(mean(xy)), absolute=quote(abs(x)), tilde=quote(X ~ Y), approx=quote(x %~~% y), equiv=quote(x %==% y), congr=quote(x %=~% y), prop=quote(x %prop% y), 'in'=quote(a %in% A), intersect=quote(intersect(A, B)), union=quote(union(A, B)), intersect=quote(is.null(intersect(A, B))))
m <- lapply(op, FUN=mathjax, flags=list(cat=FALSE))

t <- cbind(Operator=op, Output=m)
knitr::kable(t, caption="Basic elements in mathml", row.names=FALSE, escape=FALSE)
```

```{r arrows, echo=FALSE}
op <- list(twosidearrow=quote(x%<->%b), rightarrow=quote(x%->%y), leftarrow=quote(x%<-%y), uparrow=quote(x%up%y), down=quote(x%down%y), iff=quote(x%<=>%y), rArr=quote(x %=>% y), lArr=quote(x %<=% y), dblup=quote(x %dblup% y), dbldown=quote(x %dbldown% y))
m <- lapply(op, FUN=mathjax, flags=list(cat=FALSE))

t <- cbind(Operator=op, Output=m)
knitr::kable(t, caption="Arrows in mathml", row.names=FALSE, escape=FALSE)
```

@@

@@

@@

# A case study {#case-study}
